---
title: "Olympics from 1896 to 2016"
author: "Eugene Garvilles"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r global, echo=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
```

# The Olympics

## Overview

**Overview**

-   Taken from Kaggle.

-   The dataset has two files: athletes_events and regions.

-   Total of 18 columns for both datasets.

-   Height is in centimeters and weight is in kilograms.

**Required Libraries**

-   library(flexdashboard): Provides interactive dashboards using rmarkdown.

-   library(dplyr): Data manipulation library.

-   library(ggplot2): A visualization library.

## Reading in Data {data-width = 650}

**Reading in the Data**

athlete_events \<- read.csv("athlete_events.csv")

NOC_regions \<- read.csv("noc_regions.csv")

Olympics \<- merge(x = athlete_events, y = NOC_regions, by = "NOC") str(Olympics)

```{r}
athlete_events <- read.csv("athlete_events.csv")

NOC_regions <- read.csv("noc_regions.csv")

Olympics <- merge(x = athlete_events, y = NOC_regions, by = "NOC")
str(Olympics)
```

# Count Males and Females

**Count the number of Participants and Unique Participants**

Olympics %\>%

summarise(Participants = n(), Unique_Participants = n_distinct(Name))

```{r}
Olympics %>% summarise(Participants = n(), Unique_Participants = n_distinct(Name))
```

**Count the number of males and females that participated from 1896 to 2016**

Olympics %\>%

count(Sex)

```{r, echo = FALSE}
Olympics %>% 
  count(Sex)
```

# Plot Male and Female

## Column 1 {data-width = 400}

### **Figure 1. Male and Female Demographics from 1896 to 2016**

Olympics %\>%

group_by(Year, Sex) %\>%

summarise(Sports = n_distinct(Sport)) %\>%

ggplot(mapping = aes(x = Year, y = Sports, color = Sex)) +

geom_line() +

geom_smooth(se = FALSE) +

labs(x = element_blank(), y = "Sports") +

theme_minimal()+

theme(plot.title = element_text(hjust = 0.5),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank(),

panel.grid.minor.y = element_blank())

### **Figure 2. Plotting Height and Weight Between Male and Female Groups**

ann_text \<- data.frame(Height = c(175, 130, 175),

Weight = c(160, 50, 220),

Sex = c("F", "F", "M"),

label = c("this,", "This,", "and this."))

ggplot(data = Olympics, mapping = aes(x = Height, y = Weight)) +

geom_point(show.legend = FALSE, alpha = 1/10) +

facet_wrap(\~Sex) +

theme_light()+

theme(plot.title = element_text(hjust = 0.5),

panel.border = element_blank(),

panel.grid.minor.x = element_blank(),

panel.grid.minor.y = element_blank()) +

geom_text(data = ann_text,label=ann_text\$label)

## Column 2 {data-width = 600, .tabset}

### **Figure 1**

```{r, echo = FALSE}
  Olympics %>% 
  group_by(Year, Sex) %>% 
  summarise(Sports = n_distinct(Sport)) %>% 
  ggplot(mapping = aes(x = Year, y = Sports, color = Sex)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(x = element_blank(),
       y = "Sports") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
    
```

### **Figure 2**

```{r, echo = FALSE}
ann_text <- data.frame(Height = c(175, 130, 175), Weight = c(160, 50, 220), Sex = c("F", "F", "M"), label = c("this,", "This,", "and this."))

ggplot(data = Olympics, mapping = aes(x = Height, y = Weight)) +
  geom_point(show.legend = FALSE, alpha = 1/10) +
  facet_wrap(~Sex) +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  geom_text(data = ann_text,label=ann_text$label)
```

# Outliers

**Outlier 1**

Olympics %\>%

filter(Weight \>150 & Sex == "F") %\>%

select(Name, Sex, Age, Height, Year, Event, region, Medal)

```{r, echo = FALSE}
Olympics %>% 
  filter(Weight >150 & Sex == "F") %>% 
  select(Name, Sex, Age, Height, Year, Event, region, Medal)
```

**Outlier 2**

Olympics %\>%

filter(Height \< 130 & Sex == "F") %\>%

select(Name, Sex, Age, Height, Year, Event, region, Medal)

```{r, echo = FALSE}
Olympics %>% 
  filter(Height < 130 & Sex == "F") %>% 
  select(Name, Sex, Age, Height, Year, Event, region, Medal)
```

**Outlier 3**

Olympics %\>%

filter(Weight \> 200 & Sex == "M") %\>%

select(Name, Sex, Age, Height, Year, Event, region, Medal)

```{r, echo = FALSE}
Olympics %>% 
  filter(Weight > 200 & Sex == "M") %>% 
  select(Name, Sex, Age, Height, Year, Event, region, Medal)
```

# Medalists

## Column 1 {data-width = 100}

**Count the number of medals received by countries**

Olympics %\>%

filter(!is.na(Medal)) %\>%

group_by(region) %\>%

count(region, sort = TRUE)

```{r, echo = FALSE}
Olympics %>% 
  filter(!is.na(Medal)) %>% 
  group_by(region) %>% 
  count(region, sort = TRUE)
```

## Column 2 {data-width = 900}

**Plot the count of medals**

Olympics %\>%

filter(!is.na(Medal)) %\>%

filter(region %in% c("USA", "Russia", "Germany", "UK", "France")) %\>%

ggplot(mapping = aes(x = region, fill = region)) +

geom_bar(show.legend = TRUE) +

facet_wrap(\~factor(Medal, levels = c("Gold", "Silver", "Bronze"))) +

labs(title = "Medal Holders", x = "Countries", y = "Number of Medalists") +

theme_light()+

theme(plot.title = element_text(hjust = 0.5),

axis.text.x = element_blank(),

axis.ticks = element_blank(),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank(),

panel.grid.minor.y = element_blank())

```{r, echo = FALSE}
Olympics %>% 
  filter(!is.na(Medal)) %>% 
  filter(region %in% c("USA", "Russia", "Germany", "UK", "France")) %>% 
  ggplot(mapping = aes(x = region, fill = region)) +
  geom_bar(show.legend = TRUE) +
  facet_wrap(~factor(Medal, levels = c("Gold", "Silver", "Bronze"))) + 
  labs(title = "Medal Holders",
       x = "Countries",
       y = "Number of Medalists") +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

# Age

## Column 1 {data-width = 650}

### Analyzing Age

**Display statistics about age**

summary(Olympics\$Age)

```{r, echo = FALSE}
summary(Olympics$Age)
```

**Display details of 10 and 97 year-old athletes**

Olympics %\>%

filter(Age == 10 \| Age == 97) %\>%

select(Name, Sex, Age, Height, Year, Event, region, Medal)

```{r, echo = FALSE}
Olympics %>% 
  filter(Age == 10 | Age == 97) %>% 
  select(Name, Sex, Age, Height, Year, Event, region, Medal)
```

**Figure 1. Plot of Mean Age per Year from 1896 to 2016**

Olympics %\>%

filter(!is.na(Age)) %\>%

group_by(Year) %\>%

summarise(mean_Age = mean(Age)) %\>%

ggplot(mapping = aes(x = Year, y = mean_Age)) +

geom_point() +

geom_smooth() +

coord_cartesian(ylim = c(22,36)) +

labs(title = "Mean Age of All Olympics Participants from 1896 to 2016", x = "Years", y = "Mean Age") +

theme_light()+

theme(plot.title = element_text(hjust = 0.5),

axis.ticks = element_blank(),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank(),

panel.grid.minor.y = element_blank())

**Build categories to better understand age composition**

Olympics2 \<- Olympics %\>%

mutate(Age_Cat = ifelse(Age \>=10 & Age \<20, "10-19",

ifelse(Age \>=20 & Age \<30, "20-29",

ifelse(Age \>=30 & Age \<40, "30-39",

ifelse(Age \>=40 & Age \<50, "40-49",

ifelse(Age \>= 50, "Over 50", NA))))))

head(Olympics2)

```{r}
Olympics2 <- Olympics %>% 
  mutate(Age_Cat =
  ifelse(Age >=10 & Age <20, "10-19",
         ifelse(Age >=20 & Age <30, "20-29",
                ifelse(Age >=30 & Age <40, "30-39",
                       ifelse(Age >=40 & Age <50, "40-49",
                              ifelse(Age >= 50, "Over 50",
                              NA))))))

head(Olympics2)
```

**Figure 2. Plot according to Age Categories**

Olympics2 %\>%

filter(!is.na(Age)) %\>%

ggplot(mapping = aes(x = Sex, y = Age,fill = Sex)) +

geom_col() +

facet_grid(\~Age_Cat) +

labs(title = "Male and Female Distribution according to Age Category", x = element_blank(), y = element_blank()) +

theme_light()+

theme(plot.title = element_text(hjust = 0.5),

axis.text = element_blank(),

axis.ticks = element_blank(),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank(),

panel.grid.minor.y = element_blank(),

panel.grid.major.y = element_blank())

## Column 2 {data-width = 350, .tabset}

### **Figure 1**

```{r, echo = FALSE}
Olympics %>% 
  filter(!is.na(Age)) %>%
  group_by(Year) %>% 
  summarise(mean_Age = mean(Age)) %>% 
  ggplot(mapping = aes(x = Year, y = mean_Age)) +
  geom_point() +
  geom_smooth() +
  coord_cartesian(ylim = c(22,36)) +
  labs(title = "Mean Age of All Olympics Participants from 1896 to 2016",
       x = "Years",
       y = "Age") +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

### **Figure 2**

```{r, echo = FALSE}
Olympics2 %>% 
  filter(!is.na(Age)) %>% 
  ggplot(mapping = aes(x = Sex, y = Age,fill = Sex)) +
  geom_col() +
  facet_grid(~Age_Cat) +
  labs(title = "Male and Female Distribution according to Age Category",
       x = element_blank(),
       y = element_blank()) +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank())
```

# Swimming

## Column 1 {data-width = 400}

**Boxplot of Age in Swimming Sports**

Olympics %\>%

filter(!is.na(Medal) & Sport == "Swimming") %\>%

ggplot(mapping = aes(x = Medal, y = Age, fill = Medal)) +

geom_boxplot(show.legend = FALSE) +

facet_grid(\~Sex) +

labs(title = "Age of Medalists from 1896 to 2016 in Swimming", x = "Medals", y = "Age") +

theme_light() +

theme(plot.title = element_text(hjust = 0.5),

axis.ticks = element_blank(),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank())

## Column 2{data-width = 600}

```{r, echo = FALSE}
Olympics %>% 
  filter(!is.na(Medal) & Sport == "Swimming") %>% 
  ggplot(mapping = aes(x = Medal, y = Age, fill = Medal)) +
  geom_boxplot(show.legend = FALSE) +
  facet_grid(~Sex) +
  labs(title = "Age of Medalists from 1896 to 2016 in Swimming",
       x = "Medals",
       y = "Age") +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

# Gymnastics

## Column 1{data-width = 400}

**Boxplot of Age in Gymnastics Sports**

Olympics %\>%

filter(!is.na(Medal) & Sport == "Gymnastics") %\>%

ggplot(mapping = aes(x = Medal, y = Age, fill = Medal)) +

geom_boxplot(show.legend = FALSE) +

facet_grid(\~Sex) +

labs(title = "Age of Medalists from 1896 to 2016 in Gymnastics", x = "Medals", y = "Age") +

theme_light()+

theme(plot.title = element_text(hjust = 0.5),

axis.ticks = element_blank(),

panel.border = element_blank(),

panel.grid.major.x = element_blank(),

panel.grid.minor.x = element_blank())

## Column 2{data-width = 600}

```{r, echo = FALSE}
Olympics %>% 
  filter(!is.na(Medal) & Sport == "Gymnastics") %>% 
  ggplot(mapping = aes(x = Medal, y = Age, fill = Medal)) +
  geom_boxplot(show.legend = FALSE) +
  facet_grid(~Sex) +
  labs(title = "Age of Medalists from 1896 to 2016 in Gymnastics",
       x = "Medals",
       y = "Age") +
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```
